Estimating Information Transfer in Hard Parameter Sharing via Precise Bias-variance Asymptotics

### 2021-07-15 revision_journal/exp.tex
sec 4 wip

\subsection{Evaluating HPS under covariate and model shifts}

@@

\begin{wrapfigure}[12]{r}{0.3\textwidth}
	\vspace{-0.35in}
	\begin{center}
		\includegraphics[width=0.27\textwidth]{figures/mtl_model_arch.pdf}
	\end{center}
	\vspace{-0.28in}
	\caption{A hard parameter sharing architecture.}
	\label{fig_intro_arch}
\end{wrapfigure}
We conduct further studies of HPS for text classification tasks.
We consider six datasets for predicting movie review sentiment (MR and SST) \cite{pang2005seeing,socher2013recursive}, sentence subjectivity (SUBJ) \cite{pang2004sentimental}, customer review sentiment (CR) \cite{hu2004mining}, question types (TREC) \cite{li2002learning}, and polarity of a phrase (MPQA) \cite{wiebe2005annotating}.\footnote{The datasets can be downloaded at \url{https://github.com/harvardnlp/sent-conv-torch/tree/master/data}. Further statistics of each dataset can also be found following the link.}
%The question is to predict positive or negative sentiment expressed in the text.
We learn a hard parameter sharing model that consists of a word embedding layer using GloVe \cite{pennington2014glove}, followed by a shared feature representation layer.\footnote{The GloVe word vectors can be downloaded at \url{https://nlp.stanford.edu/projects/glove/}.}
A separate output layer is used for predicting the labels of each dataset.
We evaluate on three possibilities of feature representation layers including LSTM, MLP, and CNN \cite{lei2018simple}.\footnote{Before the MLP layer, we apply an average pooling layer over the embedding of every word in the sentence.}
See Figure \ref{fig_intro_arch} for an illustration of the network architecture.

@@

\begin{figure}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.8\textwidth]{figures/fig3b.pdf}
		\caption{Covariate alignment under varying sample sizes}
		\label{fig_ab_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.8\textwidth]{figures/fig3a.pdf}
		\caption{Information transfer under varying sample sizes}
		\label{fig_ab_data}
	\end{subfigure}
	\caption{Comparing hard parameter sharing (HPS) to single-task learning (STL) and a covariate alignment approach proposed by \cite{WZR20}:
	In Figure \ref{fig_ab_data}, we observe that for multiple task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually -- a phenomenon similar  to Figure \ref{fig_size}.
	This experiment fixes $n_2 = 1000$ while varying $n_1$.
	In Figure \ref{fig_ab_cov}, we observe that as task one's sample size increases, covariance alignment improves more over HPS.
	On the $y$-axis, we plot task two's test accuracy using HPS, subtracted by its STL test accuracy.
    We fix task two's sample size at $1000$ and increase task one's sample size from $100$ to $3000$.}
	\label{fig_text}
\end{figure}

%For the text classification experiment, we encode each word using the GLoVe word embeddings.%
%\textit{Predicting transfer effect via STL results.}
%We show that the single-task based metric proposed in Section \ref{sec_similarity} can predict positive or negative transfer in MTL.
%A common challenge in the study of MTL is that the results can be hard to understand.
%It is difficult to predict when MTL performs well without running extensive trials.
%Our insight is that we can use STL results to help understand MTL results.
%Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 tasks.
%We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ precision and $38.8\%$ recall among $30$ times $5$ (random seeds) task pairs!
%We observe similar results for $91$ task pairs from the ChestX-ray14 dataset.
%The results show that STL results are indicative of MTL results.

@@

\begin{wrapfigure}[9]{r}{0.3\textwidth}
	\vspace{0.0in}
	\centering
	\begin{tabular}{c c c}
		\toprule
		Models & Relative cost \\
		\midrule
		{\bf CNN}  & 30\% \\
		{\bf LSTM} & 35\% \\
		{\bf MLP}  & 31\%  \\
		\bottomrule
		\end{tabular}
		\captionof{table}{The computational cost of running Algorithm \ref{alg_inc_train} relative to the cost of running SGD.}
		\label{tab:taskonomy}
	\label{fig_intro_arch}
\end{wrapfigure}
We evaluate Algorithm \ref{alg_inc_train} under two scenarios.
First, we consider all $15$ pairwise tasks and use the first task as a data source to help predict the second task.
We find that when averaged over all $15$ two-task pairs, our procedure requires less than $35\%$ of the computational cost relative to SGD while reaching the same level of test accuracy for predicting task two.
%Our insight is that since adding more samples from the source task does not always help, we can improve efficiency by adding source samples \textit{progressively} during training.
%\textbf{Improving transfer learning training efficiency.}
%We show that Algorithm \ref{alg_inc_train} also applies to transfer learning settings.
%Compared to fine-tuning the source model on the target task, we show that our proposed method reduces the computional cost by \alert{$xx\%$}, without sacrificing accuracy.
Second, we consider using all six datasets as data sources to help predict every other task---that is, we measure the average accuracy of predicting all six tasks.
We find that a similar progressive training style procedure requires less than $35\%$ of the computational cost for reaching the same average test accuracy compared to SGD.

\medskip
\noindent\textit{Further details.} We task two's test accuracy and set $\tau$ as task two's test accuracy obtained via the standard round-robin training schedule.
We include all of task two's data and progressively add task one's data using the procedure described above.
Since the prediction accuracy has been controlled the same, we compare the computational cost.
We set $\tau$ to be the average test accuracy of all six tasks obtained using standard round-robin training.
We extend our procedure to all six datasets. We include the data from all tasks except SST. For SST, we progressively add data similar to the above procedure.

>>>>

This section complements our theoretical analysis of HPS with empirical evaluations.
First, we evaluate the performance of HPS estimators against several natural transfer learning estimators.
We show that HPS achieves superior performance under various settings of covariate and model shifts.
Second, we provide empirical implications that are inspired by our theoretical insights for mitigating covariate and model shifts.
One concerns a certain covariate alignment procedure for mitigating covariate shift.
We show that such alignment procedures provide greater gains for larger regimes of $n_1 / n_2$.
The other is a progressive training procedure for mitigating model shift.
We show that progressively increasing the sample sizes of data sources significantly reduces the computational cost of learning HPS neural networks while achieving the same accuracy for predicting the target task.

\subsection{Evaluating HPS under covariate and model shifts}\label{sec_diff}

@@

\begin{wrapfigure}[12]{r}{0.3\textwidth}
	\vspace{-0.35in}
	\begin{center}
		\includegraphics[width=0.27\textwidth]{figures/mtl_model_arch.pdf}
	\end{center}
	\vspace{-0.28in}
	\caption{A hard parameter sharing neural network.}
	\label{fig_intro_arch}
\end{wrapfigure}
We conduct further studies of HPS for text classification tasks.
We consider six datasets for predicting movie review sentiment (MR and SST) \cite{pang2005seeing,socher2013recursive}, sentence subjectivity (SUBJ) \cite{pang2004sentimental}, customer review sentiment (CR) \cite{hu2004mining}, question types (TREC) \cite{li2002learning}, and polarity of a phrase (MPQA) \cite{wiebe2005annotating}.\footnote{The datasets can be downloaded at \url{https://github.com/harvardnlp/sent-conv-torch/tree/master/data}. Further statistics of each dataset can also be found following the link.}
%The question is to predict positive or negative sentiment expressed in the text.
We learn a hard parameter sharing model that consists of a word embedding layer using GloVe \cite{pennington2014glove}, followed by a shared feature representation layer.\footnote{The GloVe word vectors can be downloaded at \url{https://nlp.stanford.edu/projects/glove/}.}
A separate output layer is used for predicting the labels of each dataset.
We evaluate on three possibilities of feature representation layers including LSTM, MLP, and CNN \cite{lei2018simple}.\footnote{Before the MLP layer, we apply an average pooling layer over the embedding of every word in the sentence.}
See Figure \ref{fig_intro_arch} for an illustration of the network architecture.

@@

\begin{figure}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.8\textwidth]{figures/fig3b.pdf}
		\caption{Covariate alignment under varying sample sizes}
		\label{fig_ab_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.8\textwidth]{figures/fig3a.pdf}
		\caption{Information transfer under varying sample sizes}
		\label{fig_ab_data}
	\end{subfigure}
	\caption{
	We extend some of our theoretical observations to text classification tasks using HPS neural networks.
	Figure \ref{fig_ab_cov} shows that covariance alignment improves the task two's test accuracy more for larger sample size ratio $n_1 / n_2$.
 	Figure \ref{fig_ab_data} shows that task one provides a positive transfer to task two in HPS only for a restricted range of $n_1$ (similar to Figure \ref{fig_sec3_model_transition}).
    Both experiments fixes $n_2 = 1000$ while varying $n_1$ from $100$ to $3000$.
	The $y$-axis measures task two's test accuracy using HPS subtracted by its test accuracy without using task one's data.}
	\label{fig_text}
\end{figure}

%\textit{Predicting transfer effect via STL results.}
%We show that the single-task based metric proposed in Section \ref{sec_similarity} can predict positive or negative transfer in MTL.
%A common challenge in the study of MTL is that the results can be hard to understand.
%It is difficult to predict when MTL performs well without running extensive trials.
%Our insight is that we can use STL results to help understand MTL results.
%Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 tasks.
%We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ precision and $38.8\%$ recall among $30$ times $5$ (random seeds) task pairs!
%We observe similar results for $91$ task pairs from the ChestX-ray14 dataset.
%The results show that STL results are indicative of MTL results.

@@

\begin{wrapfigure}[9]{r}{0.3\textwidth}
	\vspace{0.0in}
	\centering
	\begin{tabular}{c c c}
		\toprule
		Models & Relative cost \\
		\midrule
		{\bf CNN}  & 30\% \\
		{\bf LSTM} & 35\% \\
		{\bf MLP}  & 31\%  \\
		\bottomrule
		\end{tabular}
		\captionof{table}{The computational cost of running Algorithm \ref{alg_inc_train} relative to the cost of running SGD.}
		\label{tab:taskonomy}
	\label{fig_intro_arch}
\end{wrapfigure}
We evaluate Algorithm \ref{alg_inc_train} under two scenarios.
First, we consider all $15$ pairwise tasks and use the first task as a data source to help predict the second task.
We find that when averaged over all $15$ two-task pairs, our procedure requires less than $35\%$ of the computational cost relative to SGD while reaching the same level of test accuracy for predicting task two.
Second, we consider using all six datasets as data sources to help predict every other task---that is, we measure the average accuracy of predicting all six tasks.
We find that a similar progressive training style procedure requires less than $35\%$ of the computational cost for reaching the same average test accuracy compared to SGD.

\medskip
\noindent\textit{Further details.}
Algorithm \ref{alg_inc_train} uses SGD during the minimization step.
We randomly shuffle the data of both tasks and apply SGD on the shuffled data.
For the two-task experiment, we set $\tau$ to be task two's (STL) test accuracy trained without using any other task's data.
%We include all of task two's data and progressively add task one's data using the procedure described above.
As a result, once Algorithm \ref{alg_inc_train} terminates, task two's test accuracy is at least $\tau$.
For the six-task experiment, We set $\tau$ to be the average test accuracy of all six tasks trained using all tasks' data.
We include the data from all tasks except SST. For SST, we progressively increasing its sample size similar to Algorithm \ref{alg_inc_train}.

### 2021-07-15 revision_journal/results.tex
sec 4 wip

In case (i), we provide the exact asymptotics for both the bias and variance terms, together with almost sharp convergence rates. We will use these results to illustrate the effects of the \emph{bias-variance tradeoff} and \emph{varying sample sizes}. In case (ii), we will prove an exact asymptotic variance limit, and employ the result to show the intricate effect of the covariate shift. On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$.

\subsection{Covariate shift}\label{sec_diff}

@@

As a remark, when $p < n_2 < 3p$, similar results can be derived using the above arguments (details omitted).
Figure \ref{fig_sec3_model_shift} illustrates Claim \ref{claim_model_shift} for different regimes of model shift.
First, we observe that our estimates accurately match with the empirical risks.
Second, as we vary the model shift parameter $\mu$, we observe the three information transfer regimes, as Claim \ref{claim_model_shift} suggests.

\subsection{Covariate and model shift}

>>>>

We show precise estimates of the bias and variance of HPS and almost sharp convergence rates.
In Section \ref{sec3_cov}, we characterize the variance of HPS under covariate shift, which extends the classical result from Lemma \ref{fact_tr} to the sum of two sample covariance matrices.
We provide a detailed analysis of the impact of covariate shift on the excess risk of HPS.
In Section \ref{sec_sizeratio}, we characterize the bias of HPS under model shift.
Combined with the variance estimates, we describe three regimes of information transfer in the random-effect model, depending on the sample size $n_1$ and the model shift parameter $\mu$.
Finally, Section \ref{sec3_combined} extends our results to settings with both covariate and model shift.
%On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$.

\subsection{Covariate shift}\label{sec3_cov}

@@

As a remark, when $p < n_2 < 3p$, similar results can be derived using the above arguments (details omitted).
Figure \ref{fig_sec3_model_shift} illustrates Claim \ref{claim_model_shift} for different regimes of model shift.
First, we observe that our estimates accurately match with the empirical risks.
Second, as we vary the model shift parameter $\mu$, we observe the three information transfer regimes shown in Claim \ref{claim_model_shift}.

\subsection{Covariate and model shift}\label{sec3_combined}
